---
title: "Final Assignment"
author: " Pablo Romero, Irene García, Jorge Ramos"
date: "2025-04-04"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

# Introduction

Kanye West, known for his unique style and ability to fuse musical
genres, has sparked numerous debates both for his music and his behavior
offstage. Recently, he has been at the center of media attention due to
controversial remarks and attitudes, including neo-Nazi comments and his
contentious treatment of his partners (such as his ex-wife, Kim
Kardashian, or his current wife, Bianca Censori). These situations have
raised growing concerns about the potential impact of his persona on
society.

Text analysis is a fundamental tool for this purpose, allowing us to
extract patterns and insights from large volumes of textual data. To
achieve this, three different techniques have been selected:

1.  Sentiment Analysis: The objective is to explore the emotions and
    sentiments expressed in his songs throughout his career.

2.  Term Frequency - TF-IDF: The objective is to detect rare words in
    albums compared to others and see how representative offensive words
    are.

3.  

```{r setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = TRUE)
```

```{r}
set.seed(123)
rm(list = ls())
```

# Packages

```{r libraries}
library(dplyr)
library(readr)
library(ggplot2)
library(tidyverse)
library(geniusr)
library(tidytext)
library(quanteda)
library(devtools)
library(forcats)
library(wordcloud)
library(RColorBrewer)
library(ggraph)
library(igraph)
library(widyr)
```

# Download the data

```{r}
# Read the data from GitHub
kanye_west <- read.csv("https://github.com/babyakja/GA_capstone_project/raw/master/data/Kanye%25West.csv")
```

The selected dataframe comes from a GitHub repository containing a total
of 98 Kanye West songs, covering his discography from 2004 to 2018.
However, the original CSV file does not include album identification or
release years, so this information will need to be added manually to
complete the analysis.

```{r}
# Create a vector including the songs for each album
album_data <- list(
  "The College Dropout (2004)" = c("We Don't Care", "Graduation Day", "All Falls Down", "Spaceship", "Jesus Walks", "Never Let Me Down", 
                                   "Get Em High", "Workout Plan", "The New Workout Plan", "Slow Jamz", "Breathe In Breathe Out",
                                   "School Spirit", "Lil Jimmy Skit", "Two Words", "Through the Wire", "Family Business", "Last Call", "I'll Fly Away"),
  
  "Late Registration (2005)" = c("Wake Up Mr. West", "Heard 'Em Say", "Touch the Sky", "Gold Digger", "Drive Slow", "My Way Home", 
                                 "Crack Music", "Roses", "Bring Me Down", "Addiction", "Diamonds From Sierra Leone (Remix)", 
                                 "We Major", "Hey Mama", "Celebration", "Gone", "Late"),
  
  "Graduation (2007)" = c("Good Morning", "Champion", "Stronger", "I Wonder", "Good Life", "Can't Tell Me Nothing", "Barry Bonds", 
                          "Drunk and Hot Girls", "Flashing Lights", "Everything I Am", "The Glory", "Homecoming", 
                          "Big Brother"),
  
  "808s & Heartbreak (2008)" = c("Say You Will", "Welcome To Heartbreak", "Heartless", "Amazing", "Love Lockdown", "Paranoid", 
                                 "RoboCop", "Street Lights", "Bad News", "See You in My Nightmares", "Coldest Winter"),
  
  "My Beautiful Dark Twisted Fantasy (2010)" = c("Dark Fantasy", "Gorgeous", "Power", "All Of The Lights (Interlude)", "All of the Lights", 
                                                "Monster", "So Appalled", "Devil in a New Dress", "Runaway", "Hell of a Life", 
                                                "Blame Game", "Lost in the World", "Who Will Survive In America"),
  
  "Yeezus (2013)" = c("On Sight", "Black Skinhead", "I Am a God", "New Slaves", "Blood On The Leaves", "Bound 2"),
  
  "The Life of Pablo (2016)" = c("Famous", "Feedback", "Low Lights", "Highlights", "Freestyle 4", "I Love Kanye", "Wolves", 
                                 "Facts (Charlie Heat Version)", "Fade", "Saint Pablo"),
  
  "Kids See Ghosts (2018)" = c("Kids See Ghosts", "Freeee (Ghost Town, Pt. 2)", "Fire", "4th Dimension", "Reborn", "Feel the Love", "Cudi Montage"),
  
  "Ye (2018)" = c("I Thought About Killing You", "Yikes", "All Mine", "Wouldn't Leave", "No Mistakes", "Ghost Town", "Violent Crimes")
)

# We create a new column and introduce the new info
kanye_west$album <- sapply(kanye_west$name, function(x) {
  album_found <- NA
  for (album in names(album_data)) {
    if (x %in% album_data[[album]]) {
      album_found <- album
      break
    }
  }
  return(album_found)
})

# We use the information between parentheses to extract the release year
kanye_west <- kanye_west %>%
  mutate(
    release_year = str_extract(album, "\\((\\d{4})\\)") %>% str_remove_all("[()]") %>% as.integer()
  )

# Finally, we convert the df into a tibble and verify the first columns
kanye_west <- as_tibble(kanye_west)
head(kanye_west)

# We check for NAs
kanye_west %>% 
  filter(is.na(album) | is.na(release_year))
```

With this addition, we now have the necessary information to carry out
the analysis. As observed, there are a total of 9 albums and one song
that does not belong to any, meaning it is a single. We don't know why
it was included in the dataset, because it is not the only single
released by the artist in this time period. Since we're interesed in
working at the album level, we will drop this observation now:

```{r}
kanye_west <- kanye_west %>%
  filter(!is.na(album))
```

The next step is to clean the data set: In the text variable, there is
extra information that does not correspond to the lyrics but instead
includes details about the production or who sings in specific parts of
the song (e.g., Intro, Verse 1, etc.). If this information were
available for all songs, analyzing the producers or collaborators could
provide valuable insights to enrich the study. However, since it is not
consistently present across all tracks, it is best to remove it to avoid
potential issues.

```{r}
# Remove the brackets and the information inside
kanye_west$text <- gsub("\\[.*?\\]", "", kanye_west$text)
```

# Tokenization and stopwords:

The lyrics of Kanye West's songs will be tokenized to make working with
words more manageable and structured.

1.Punctuation will be removed since, in a text analysis like this, it
does not add value to identifying key words or term frequency.

2.All words will be converted to lowercase to prevent the same words in
uppercase and lowercase from being treated as different.

```{r}
# Tokenization with punctuation removal 
kanye_tokens <- kanye_west %>%
  unnest_tokens(word, text, drop = TRUE, strip_punct = TRUE, to_lower = TRUE) # drop = TRUE to remove the text column  

head(kanye_tokens)
```

```{r}
kanye_tokens %>%
  group_by(album, release_year) %>%
  summarise(
    number_of_songs = n_distinct(name),
    number_of_words = n(),
    .groups = "drop"
  ) %>%
  arrange(release_year)
```

Now, we filter the stopwords (from `tidytext package`) in order to
eliminate those uninformative words:

```{r}
kanye_filtered <- kanye_tokens %>%
  anti_join(stop_words)
```

With these steps, we now have our table ready for analysis, ensuring
that the data is clean and structured.

# Introductory Analysis

Before delving into the selected techniques, it is necessary to conduct
an introductory analysis to better understand the distribution and
characteristics of the words present in Kanye West's lyrics.

```{r}
# Count the frequency of words
word_counts <- kanye_filtered %>%
  count(word, sort = TRUE)

# top 10 most frequent words
head(word_counts, 10)
```

From this table we can already get an idea of the preeminence of swear
words, slurs and slang in Kanye's discography, but we will have to
conduct a deeper analysis to get a more accurate picture.

The top word, "la", doesn't seem to be very informative, but let's see
it in context to have a better idea:

```{r}
corpus <- corpus(kanye_west$text)

kwic(tokens(corpus), "la", window = 5)
```

The word 'la', appears to carry three different meanings in Kanye's
songs: an article in a foreign language, LA (Los Angeles) and, most
commonly, a vocable - a wordless sound or syllable with no semantic
meaning used in music.

Let's make a plot and a wordcloud to see this more visually:

```{r}
album_order <- c(
  "The College Dropout (2004)", 
  "Late Registration (2005)", 
  "Graduation (2007)", 
  "808s & Heartbreak (2008)", 
  "My Beautiful Dark Twisted Fantasy (2010)", 
  "Yeezus (2013)", 
  "The Life of Pablo (2016)", 
  "Kids See Ghosts (2018)", 
  "Ye (2018)"
)
# Visualization of words frequency
word_counts %>%
  filter(n > 50) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(x = word, y = n, fill = n)) + 
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() +  
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  labs(title = "Most frequent words in Kanye West lyrics",
       x = NULL,
       y = "Frequency") +
  theme_minimal()

# Generate the wordcloud
wordcloud(words = word_counts$word, 
          freq = word_counts$n, 
          min.freq = 5, 
          max.words = 200,  
          scale = c(4, 0.5),  
          colors = brewer.pal(8, "Dark2"), 
          random.order = FALSE)
```

Both of these expand on the idea from the previous table: we have some
seemingly meaningless syllables, similar to "la", such as "yeah", "hey",
"uh" or "bam", as well as some more examples of swearwords and slang:
"fuck", "ass", "damn", "yo"...

We verify the presence of words referring to women, which we could also
anticipate: "girl", "baby", "mama", "hoes", "bitch"...

Some of the words in the graphs, like "time", "God", "money" or "kids",
might reveal other topics present in Kanye's discography; but, in
general, we can say that the most common words are pretty ambiguous, and
can take on very different meanings depending on the context, for which
we will have to take our analysis some steps further.

The following chart shows the 10 most common words per album:

```{r}
#Visualization:

kanye_filtered %>%
  count(album, release_year, word, sort = TRUE) %>%
  group_by(album) %>%
  top_n(10, n) %>%
  ungroup() %>%
  mutate(album = factor(album, levels = album_order),  # Orders albums chronologically
         word = reorder_within(word, n, album)) %>%
  ggplot(aes(n, word, fill = album)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ album, scales = "free_y", ncol = 5) +
  scale_y_reordered() +
  labs(title = "Most frequent words per album",
       x = "Count",
       y = NULL) +
  theme_minimal() +
  theme(strip.text = element_text(size = 8))


```

These plots represent the words that appear most often in each album,
regardless of album length, so albums with more or longer songs may
dominate with higher raw counts.

As can be seen in the graphs, there are words that appear across several
albums by Kanye West, such as "feel", "girl", "love", or "shit". Most
albums don't have very prominent words, with most words on the top of
the ranking having very similar counts.

Next, it's useful to replicate this plot using *term frequency* instead
of word counts. This implies normalizing the count of each word by the
total number of words in that album, which is more insightful when
comparing across albums of different lengths.

```{r}
kanye_filtered %>%
  count(album, release_year, word) %>%
  group_by(album) %>%
  mutate(total_words = sum(n),
         tf = n / total_words) %>%
  top_n(10, tf) %>%
  ungroup() %>%
  mutate(album = factor(album, levels = album_order),
         word = reorder_within(word, tf, album)) %>%
  ggplot(aes(tf, word, fill = album)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ album, scales = "free_y", ncol = 5) +
  scale_y_reordered() +
  labs(title = "Top 10 most characteristic words per album (Term Frequency)",
       x = "Term Frequency",
       y = NULL) +
  theme_minimal() +
  theme(strip.text = element_text(size = 8))
```

For College Dropout and My Dark Twisted Fantasy specially, we can see
that the frequency of the top words considering album length is very
low, which implies that the vocabulary of those albums is probably more
varied and balanced.

On the other hand, the higher TF can be found on Kids See Ghosts, which
is likely due to the fact that we only have two songs for this album.

Lastly, we can plot the words with the highest TF-IDF, which will allow
us to see what the most distinctive words for each album inside the
collection are.

```{r}
kanye_filtered %>%
  count(album, release_year, word, sort = TRUE) %>%
  bind_tf_idf(term = word, document = album, n = n) %>%
  group_by(album) %>%
  slice_max(tf_idf, n = 10) %>%
  ungroup() %>%
  mutate(
    album = factor(album, levels = album_order),
    word = reorder_within(word, tf_idf, album)
  ) %>%
  ggplot(aes(tf_idf, word, fill = album)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ album, scales = "free", ncol = 5) +
  scale_y_reordered() +
  labs(
    title = "Top 10 most distinctive words per album (TF-IDF)",
    x = "TF-IDF score",
    y = NULL
  ) +
  theme_minimal() +
  theme(strip.text = element_text(size = 8))
```

------- no sé si todos estos gráficos tendríamos que haberlos hecho con lemmas
en vez de con las palabras tal cual......
aquí dejo el código del gráfico de TF-IDF con lemmas, aunque el objeto de 
tfidf_all lo creamos más tarde

```{r}
# TF-IDF WITH LEMMAS INSTEAD OF ORIGINAL WORDS
# tfidf_all %>%
#   group_by(album) %>%
#   slice_max(tf_idf, n = 10) %>%
#   ungroup() %>%
#   mutate(album = factor(album, levels = album_order),
#     word_lemma = reorder_within(word_lemma, tf_idf, album)  # Using word_lemma in place of 'word'
#   ) %>%
#   ggplot(aes(tf_idf, word_lemma, fill = album)) +
#   geom_col(show.legend = FALSE) +
#   facet_wrap(~ album, scales = "free", ncol = 5) +
#   scale_y_reordered() +
#   labs(
#     title = "Top 10 Most Distinctive Words per Album (TF-IDF)",
#     x = "TF-IDF Score",
#     y = NULL
#   ) +
#   theme_minimal() +
#   theme(strip.text = element_text(size = 8))
```

By visualizing the TF-IDF, each album reveals a unique character, giving
us a better insight into what distinctive topics Kanye might be covering
in each of them. The terms displayed are not only frequently used in the
respective albums but are also significant in distinguishing these
albums from others.

However, we must bear in mind that, for most bars, TF-IDF scores are
pretty low. This may point to a significant vocabulary overlap: Kanye
often reuses similar words or themes across all his albums.

Finally, the proportion of each word per album will be calculated, and
then the correlation between albums will be analyzed. In this case, two
albums will be selected:

```{r}
# First, the proportion:
frequency <- kanye_filtered %>%
  count(album, word) %>%
  group_by(album) %>%
  mutate(proportion = n / sum(n)) %>%
  ungroup() %>%
  pivot_wider(names_from = album, values_from = proportion, values_fill = 0)

# Now the correlations:
cor.test(data = frequency, ~ `Graduation (2007)` + `My Beautiful Dark Twisted Fantasy (2010)`)

cor.test(data = frequency, ~ `Yeezus (2013)` + `Ye (2018)`)
```

n the first analysis, the correlation between Graduation (2007) and My
Beautiful Dark Twisted Fantasy (2010) is very weak and negative
(-0.0448), with a significant p-value (0.0004), suggesting a non-zero
correlation.

In the second analysis, the correlation between Yeezus (2013) and Ye
(2018) is nearly zero (-0.0022), with a high p-value (0.862), indicating
no significant correlation.

# Sentiment Analysis

The goal of this analysis is to understand the emotions and tone of
Kanye West's lyrics throughout his discography. To achieve this, we will
apply three complementary sentiment analysis approaches:

1.NRC: This allows us to identify specific feelings (joy, sadness,
anger, fear, etc.) in his songs, providing a more nuanced view of his
lyrics.

2.Bing: To classify words as positive and negative, giving us an overall
sense of the sentiment of each album or song.

3.AFINN: To apply a numerical scale to the sentiment of words, allowing
us to quantify the overall tone of each song and observe changes in
intensity over time.

To carry out the sentiment analysis, we first need to load the lexicons
and apply them to our data:

```{r}
# Load the lexicons
nrc_lexicon <- get_sentiments("nrc")
bing_lexicon <- get_sentiments("bing")
afinn_lexicon <- get_sentiments("afinn")

# Merge the dataframe with the lexicons
kanye_nrc <- kanye_filtered %>%
  inner_join(nrc_lexicon, by = "word")

kanye_bing <- kanye_filtered %>%
  inner_join(bing_lexicon, by = "word")

kanye_afinn <- kanye_filtered %>%
  inner_join(afinn_lexicon, by = "word")

```

### 1. Visualización de emociones con NRC

```{r}
kanye_nrc %>%
  count(sentiment, sort = TRUE) %>%
  ggplot(aes(x = reorder(sentiment, n), y = n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(title = "Distribution of emotions in Kanye's lyrics",
       x = "Emotion",
       y = "Frequency") +
  theme_minimal()

# Crear tabla con el conteo de emociones
emotion_table <- kanye_nrc %>%
  count(sentiment, sort = TRUE)

# Ver la tabla
emotion_table
```

The lyrics reflect a wide range of emotions, with a predominance of
positive and negative feelings, which are common in most of his songs.
Emotions such as trust, joy, and anger also appear frequently,
suggesting that the artist expresses a broad spectrum of feelings in his
music, from euphoria to frustration and reflection.

For a deeper analysis, the following graph shows the 4 most frequent
emotions in the lyrics by album:

```{r}
kanye_nrc %>%
  group_by(album, release_year, sentiment) %>% 
  count() %>%  
  group_by(album) %>% 
  slice_max(n, n = 4) %>%  
  ungroup() %>%
  mutate(
    album = factor(album, levels = album_order),
    sentiment = reorder_within(sentiment, n, album)
  ) %>%
  ggplot(aes(x = n, y = sentiment, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ album, scales = "free", ncol = 5) +
  scale_y_reordered() +
  labs(title = "4 most frequent emotions per album",
       x = "Frequency",
       y = NULL) +
  theme_minimal() +
  theme(strip.text = element_text(size = 8))
```

In this analysis, the most frequent emotions per album of Kanye West are
observed. Throughout his career, predominant sentiments include both
positive and negative feelings, reflecting the emotional complexity in
his lyrics. Additionally, sentiments such as joy, anger, and trust
regularly stand out in several albums, suggesting that Kanye explores a
wide range of emotional states, from euphoria to frustration and
distrust.

For example, in the album 808s & Heartbreak (2008), the sentiments
negative and positive are the most prominent, while in Graduation
(2007), there is a strong presence of both positive and negative
sentiments along with anger and trust. In contrast, in Kids See Ghosts
(2018), sentiments are less frequent, but positive and negative
sentiments still stand out.

### 2. Classification with BING:

What will be done is a binary classification between words considered
positive and negative.

```{r}
kanye_bing %>%
  count(sentiment) %>%
  ggplot(aes(x = sentiment, y = n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  labs(title = "Count of Positive and Negative Words",
       x = "Sentiment",
       y = "Frequency") +
  theme_minimal()
```

With Bing's binary classification, it is clear that there is a
predominance of words considered negative. As observed earlier, some of
the most frequent words were "fuck" and "nigga," so this result appears
to be consistent with the previous findings.

Now, let's examine how this classification looks by album:

```{r}
kanye_bing %>%
  count(album, release_year, sentiment) %>%  
  mutate(album = factor(album, levels = album_order)) %>%  
  ggplot(aes(x = sentiment, y = n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ album, scales = "free", ncol = 5) +  
  labs(title = "Number of Positive and Negative Words by Album",
       x = "Sentiment",
       y = "Frequency") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.text = element_text(size = 8))   

```

Similar to the previous chart, in all albums except for one (Kids See
Ghosts from 2018), there are more words classified as negative than
positive. However, the gap appears to be larger in more recent albums
(except for the aforementioned one), suggesting a trend towards the use
of negative words.

### 3. Evolution of sentiement with AFINN:

In this part, we will analyze the evolution of sentiment using the AFINN
lexicon scores (from -3 to 3):

```{r}
kanye_afinn %>%
  group_by(album, release_year) %>%
  summarise(sentiment_score = sum(value), .groups = "drop") %>%
  mutate(album = factor(album, levels = album_order)) %>%
  ggplot(aes(x = album, y = sentiment_score, fill = sentiment_score)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(title = "Sentiment Evolution in Kanye's Albums",
       x = "Album",
       y = "Sentiment Score") +
  theme_minimal() +
  theme(strip.text = element_text(size = 8))

```

In general, it stands out that the vast majority of albums have a
negative score, which seems to confirm what we saw earlier—the large
presence of negative words in the songs, and specifically, words with a
high negative score.

On the other hand, we see that there is a positive total score in only
three cases:

808 & Heartbreak (2008), which has the highest score. This is curious
because in this case, the frequency of negative words was greater than
the positive ones, so it seems that the positive words present have a
high score, or the classification was different. Kids See Ghosts (2018)
and the Single also have a positive score, although the result is
tighter.

What's been done earlier is in absolute terms, which can distort the
reality since some albums contain more songs than others. Therefore, we
will also perform a relative analysis using the average instead of the
sum of scores:

```{r}
kanye_afinn %>%
  group_by(album, release_year) %>%
  summarise(sentiment_score_avg = mean(value)) %>% 
  mutate(album = factor(album, levels = album_order)) %>%
  ggplot(aes(x = album, y = sentiment_score_avg, fill = sentiment_score_avg)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(title = "Sentiment Evolution in Kanye's Albums",
       x = "Album",
       y = "Sentiment Score") +
  theme_minimal() +
  theme(strip.text = element_text(size = 8))

```

Now, the same trend is observed: the majority of albums have a negative
average score, with the three previously mentioned albums still
maintaining a positive score. However, since these are average scores,
some changes are noticeable. 808s & Heartbreak has the highest positive
average score, and Yeezus (2013) is the album with the lowest average
score.

The sentiment analysis of Kanye West's lyrics reveals a complex
emotional landscape throughout his discography. Across his albums, there
is a clear predominance of both positive and negative feelings,
reflecting the emotional depth and diversity in Kanye's music. Emotions
such as joy, anger, and trust appear regularly, suggesting a broad
spectrum of emotional exploration, ranging from euphoria to frustration
and introspection.

The analysis also highlights a consistent presence of negative
sentiment, particularly in more recent albums. Despite the strong
negative presence, certain albums like 808s & Heartbreak (2008) and Kids
See Ghosts (2018) show a mix of both positive and negative emotions,
with a tendency towards more positive sentiment in some cases. The
evolution of sentiment over time demonstrates the emotional shifts in
Kanye’s music, with some albums leaning more heavily on negative tones,
while others explore a balance of emotions.

It’s also worth mentioning that the style of music Kanye West creates
often involves the use of slang, colloquial expressions, and insults,
which further adds to the emotional intensity of his lyrics.

#### INSULT ANALYSIS???

To analyze this, we will use the **Luis von Ahn List (LDNOOBW)**, which
is a database of offensive words in English. It includes insults and
slang commonly used in hip-hop. This list is publicly available on
GitHub.

What we will do is first examine the relative frequency of insults by
album to enable comparison across Kanye's discography. Then, we will
conduct an AFINN sentiment analysis to determine whether albums with a
higher proportion of insults tend to have more negative sentiment
scores.

```{r}
# First, we obtain the list of insults from GitHub:
url <- "https://raw.githubusercontent.com/LDNOOBW/List-of-Dirty-Naughty-Obscene-and-Otherwise-Bad-Words/master/en"
bad_words_vonahn <- readLines(url)


# Then, we calculate the total number of words per album
total_words_by_album <- kanye_filtered %>%
  group_by(album) %>%
  summarise(total_words = n())

# We lemmatize the offensive words 
kanye_offensive <- kanye_filtered %>%
  filter(word %in% bad_words_vonahn)

# Also we need to calculate the insults per album
insults_by_album <- kanye_offensive %>%
  group_by(album) %>%
  summarise(
    insult_count = n()
  ) |> 
  left_join(total_words_by_album, by = "album") |> 
  mutate(relative_insult_freq = insult_count / total_words)

# then, the visualization:
ggplot(insults_by_album, aes(x = reorder(album, relative_insult_freq), y = relative_insult_freq, fill = relative_insult_freq)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(title = "Relative Frequency of Insults by Album",
       x = "Album",
       y = "Relative Frequency of Insults") +
  theme_minimal()
```

As can be seen, the albums *Yeezus* (2013) and *My Beautiful Dark
Twisted Fantasy* (2010), which had the lowest sentiment scores in the
previous AFINN analysis, also stand out in terms of the highest
frequency of insults. Similarly, the albums with positive sentiment
scores appear to be the ones with the least number of insults.

For a deeper analysis, we will cross the number of insults with the
AFINN sentiment score to see if albums with more insults tend to have
more negative sentiment.

```{r}
# First, we get the average sentiment scores per album using AFINN 
sentiment_by_album <- kanye_afinn %>%
  group_by(album) %>%
  summarise(sentiment_score_avg = mean(value))  # Average sentiment score

insults_by_album <- insults_by_album |> 
  left_join(sentiment_by_album, by = "album") # join to the existing dataframe

# Relationship between relative insult frequency and average sentiment score
ggplot(insults_by_album, aes(x = relative_insult_freq, y = sentiment_score_avg, label = album)) +
  geom_point(color = "red", size = 3) +  # albums
  geom_smooth(method = "lm", se = FALSE, color = "blue") +  # line
  geom_text(vjust = -0.5, size = 4) +  # Album labels
  labs(title = "Relationship Between Relative Insult Frequency and Average Sentiments",
       x = "Relative Frequency of Insults",
       y = "Average AFINN Sentiment Score") +
  theme_minimal()

# Finally, the correlation to confirm the relationship
correlation_result <- cor(insults_by_album$relative_insult_freq, insults_by_album$sentiment_score_avg)
correlation_result
```

This graph and the correlation confirm this trend, meaning that albums
with a higher number of swear words generally have lower or more
negative sentiment scores. As mentioned, the two albums Yeezus and My
Beautiful Dark Twisted Fantasy have a higher frequency of curse words
and, at the same time, lower average AFINN sentiment scores. On the
other hand, 808s & Heartbreak, for example, has a positive AFINN score
and a low presence of insults.

```{r}
range(insults_by_album$relative_insult_freq)
mean(insults_by_album$relative_insult_freq)
```

The relative frequency of swear words ranges from 0.002 to 0.087. On
average, Kanye's lyrics contain less than 4 swear words for every 100
words.

Lastly, we're going to visualize the most common insults or bad words in
Kanye’s lyrics in a wordcloud. For this purpose, we must first lemmatize
the words in our dataset, so that their variations (plurals, verb forms,
etc.) are not considered as different words.

```{r}
kanye_offensive <- kanye_offensive |> 
  mutate(
         # word_stem = SnowballC::wordStem(word),
         word_lemma = textstem::lemmatize_words(word)
    )
```

Now, we can check how many unique words we have and how many unique
lemmas were created:

```{r}
kanye_offensive |>
  summarise(
    total_words = n(),
    unique_words = n_distinct(word),
    unique_lemmas = n_distinct(word_lemma)
  )

# what lemmas are there?
unique(kanye_offensive$word_lemma)
```

We have 36 words and 33 unique lemmas, which means that only 3 words
were actually replaced by their canonical form through lemmatization.

We can run a check to see what lemmas were created. We observe that some
words weren't correctly lemmatized, this is possibly because insults and
swearwords do not appear in the dictionaries. We can manually assign
these words to the most suitable lemma:

```{r}
kanye_offensive <- kanye_offensive %>%
  mutate(
    word_lemma = case_when(
      word == "titties" ~ "titty",
      word == "fuckin"  ~ "fucking",
      TRUE              ~ word_lemma
    )
  )
```

The last step is to plot the wordcloud:

```{r}
# Count the frequency of bad words 
insult_word_counts <- kanye_offensive %>%
  count(word_lemma, sort = TRUE)

# Create a word cloud for the most common bad words
set.seed(123)
wordcloud(words = insult_word_counts$word_lemma, 
          freq = insult_word_counts$n, 
          min.freq = 1, 
          max.words = 100, 
          scale = c(5, 2), 
          colors = brewer.pal(8, "Dark2"))

```

The most common insults in Kanye West's lyrics are "shit", "nigga",
"fuck", "ass", "bitch", "motherfucker" and "dick", which we believe is a
common trend in American hip-hop by male artists.

TF-IDF might serve us to analyze the representativeness of offensive
words, although we anticipate that the scores will be fairly low, since
we know that there is a short list of swear words which are likely to
appear several time across all of Kanye's albums.

```{r}
# Calculate TF-IDF for all our words:

kanye_filtered <- kanye_filtered |> 
  mutate(word_lemma = textstem::lemmatize_words(word))

tfidf_all <- kanye_filtered %>%
  count(album, word_lemma, sort = TRUE) %>%
  bind_tf_idf(word_lemma, album, n) %>%
  arrange(desc(tf_idf))

# Then, filter the offensive words
tfidf_offensive <- tfidf_all %>%
  filter(word_lemma %in% bad_words_vonahn) 

# Visualize the most important offensive words by TF-IDF for each album:

tfidf_offensive %>%
  group_by(album) %>%
  slice_max(tf_idf, n = 10) %>%
  ungroup() %>%
  mutate(album = factor(album, levels = album_order),
    word_lemma = reorder_within(word_lemma, tf_idf, album)
  ) %>%
  ggplot(aes(tf_idf, word_lemma, fill = album)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ album, scales = "free", ncol = 5) +
  scale_y_reordered() +
  labs(
    title = "Top 10 most distinctive swear words per Album (TF-IDF)",
    x = "TF-IDF score",
    y = NULL
  ) +
  theme_minimal() +
  theme(strip.text = element_text(size = 8))
```

As we were expecting, the scores are low and most curse words with the
top TF-IDF scores are repeated across the albums. This makes sense
because swear words are usually not that descriptive or meaningful, but
rather complements, attached to other words for emphasis, to express
aggressiveness, etc. Due to their nature and the relatively low number
of distinct swear words found in Kanye's lyrics, many of them are common
to all of his albums and not distinctive of any of them.

# Word co-occurences

For this exercise, we will analyze the word co-occurences in Kanye's
songs at the line level.

The first thing we need to do is separate the lines from our original
dataset, which we had called "kanye_west". We can split the text by the
line breaks, which are signalled as "\n":

```{r}
lines_df <- kanye_west %>%
  mutate(line = str_split(text, pattern = "\n")) %>%
  unnest(line) %>%
  filter(str_detect(line, "\\w")) |> # remove empty lines
  mutate(line_id = row_number())
```

We can now tokenize words keeping the line id, and then count how many
times words co-appear in the same line, using the pairwise_count function:
```{r}
words_by_line <- lines_df %>%
  unnest_tokens(word, line) |> 
  anti_join(stop_words)

word_pairs <- words_by_line %>%
  pairwise_count(item = word, feature = line_id, sort = TRUE)
```

It might be interesting to analyze, for example, what words appear
together with words referring to women in Kanye's lyrics:
```{r}
word_pairs %>%
  filter(item1 %in% c("girl", "bitch", "baby", "girls", "bitches"),
         n > 2) %>%
  arrange(desc(n))
```

FALTA TEXTO